---
title: "Transfusion Service Analysis"
author: "Jessica Jones"
date: "`r format(Sys.Date(),'%B %d, %Y')`"
output: 
  flexdashboard::flex_dashboard:
  orientation: rows
html_document:
  df_print: paged
---
  
```{r setup, include = FALSE}
library(tidyverse)
library(lubridate)
library(ggplot2)
library(plotly)
library(flexdashboard)

##most recent year
file_id <- as.numeric(format(Sys.Date(), '%Y'))-1

##variables
bbarea <- c("BB", "C", "CRC", "I", "R")
labloc <- c("B", "C", "CRC", "I", "R")
products <- c("RC", "PLTG", "PLG", "CRYG", "MFP")
product_labels <- c("RBCs", "Platelets", "Plasma", "Cryoprecipitate", "Manufactured")

## formatting
mycolors <- c("#E8000D", "#0051BA", "#FFC82D", "#355C7D", "#85898A")

Graph_Theme <- theme(plot.title = element_text(face = "bold", size = (12), hjust = 0.5),
                     plot.subtitle = element_text(size = (8), hjust = 0.5),
                     axis.title = element_text(face = "plain", size = 10),
                     plot.caption = element_text(hjust = 1),
                     legend.text = element_text(size = 8),
                     legend.key.size =  unit(5,"mm"),
                     #legend.position = "none",
                     axis.text = element_text(size = 8))

```

```{r data}
##read data
##query in tsa_query
bb_if_data <- readRDS(file = paste0("../Final/Data/bb_if_data_fy", file_id, ".rds"))
bb_is_data <- readRDS(file = paste0("../Final/Data/bb_is_data_fy", file_id, ".rds"))
bb_t_data <- readRDS(file = paste0("../Final/Data/bb_t_data_fy", file_id, ".rds"))

##data prep
bb_if_data <- bb_if_data |> 
  filter(UnitArea %in% bbarea) |>
  mutate(Month = month(IssueDate_ODBC, label = TRUE),
         Weekday = wday(IssueDate_ODBC, label = TRUE),
         Components = factor(CompGroup, 
                            levels = products,
                            labels = product_labels))

bb_is_data <- bb_is_data |> 
  filter(UnitArea %in% bbarea) |>
  mutate(Month = month(IssueDate, label = TRUE),
         Weekday = wday(IssueDate, label = TRUE),
         Hour = hour(IssueTime),
         Components = factor(CompGroup, 
                            levels = products,
                            labels = product_labels))

bb_t_data <- bb_t_data |>
  filter(ResultLab %in% labloc,
         CurrentResult == "Yes",
         Result != "HIDE") |> 
  mutate(Month = month(ReceiveDateODBC, label = TRUE),
         Weekday = wday(ReceiveDateODBC, label = TRUE),
         Hour = hour(ReceiveTimeODBC),
         Priority = case_when(PriorityCodes == "S" ~ "Stat",
                              TRUE ~ "Routine"),
         TAT = round(as.numeric(difftime(as.POSIXct(paste(ResultDateODBC, ResultTimeODBC)),
                                         as.POSIXct(paste(ReceiveDateODBC, ReceiveTimeODBC)),
                                         units = c("mins")))),
         Battery = case_when(BatTstCode == "XM" ~ "Crossmatch",
                             BatTstCode == "TYSC" ~ "Type and Screen",
                             TRUE ~ "Other %AS battery"),
         Battery = factor(Battery, levels = c("Crossmatch", "Type and Screen", "Other %AS battery")),
         Results = case_when(!(is.na(ResultModifiers)) ~ "Requires Follow up",
                             Result == "%NEG" ~ "Negative",
                             TRUE ~ "Requires Follow up"),
         Method = case_when(ResultMethod == "%BPW" ~ "Bench",
                            ResultMethod == "EC2" ~ "Echo",
                            ResultMethod == "NE1" ~ "Neo",
                            TRUE ~ "Other")) |> 
  filter(TAT > 24 & TAT < 1440)   ##remove data not representative of normal bb processes

```

```{r}
if(min(bb_is_data$IssueDate) == min(bb_if_data$IssueDate_ODBC) & min(bb_t_data$ReceiveDateODBC) == min(bb_is_data$IssueDate)) {
  mindate <- format(min(bb_is_data$IssueDate), "%B %Y")
} else {
  stop("Check report start dates.")
}

if(max(bb_is_data$IssueDate) == max(bb_if_data$IssueDate_ODBC) & max(bb_t_data$ReceiveDateODBC) == max(bb_is_data$IssueDate)) {
  maxdate <- format(max(bb_is_data$IssueDate), "%B %Y")
} else {
  stop("Check report end dates.")
}

if(min(year(bb_is_data$IssueDate)) != max(year(bb_is_data$IssueDate))){
  fiscal_months <- c("Jul", "Aug",  "Sep",  "Oct",  "Nov",  "Dec", "Jan",  "Feb",  "Mar",  "Apr",  "May",  "Jun")
  bb_if_data <- bb_if_data |> 
    mutate(Month = factor(Month, levels = fiscal_months))
  bb_is_data <- bb_is_data |> 
    mutate(Month = factor(Month, levels = fiscal_months))
  bb_t_data <- bb_t_data |> 
    mutate(Month = factor(Month, levels = fiscal_months))
}

```

Closer look at the Transfusion Service from `r mindate` to `r maxdate`.

# Monthly Overview

```{r}
##monthly counts
monthly_if_count <- bb_if_data |>
  group_by(Month) |> 
  count(name = "Transfused")

bb_if_monthlyavg <- round(mean(monthly_if_count$Transfused))

monthly_is_count <-  bb_is_data |> 
    group_by(Month) |> 
    count(name = "Issued")

bb_is_monthlyavg <- round(mean(monthly_is_count$Issued))

monthly_t_count <- bb_t_data |>  
  group_by(Month) |> 
  count(name = "Antibody Screens")

bb_t_monthlyavg <- round(mean(monthly_t_count$`Antibody Screens`))

if(nrow(monthly_is_count) == nrow(monthly_if_count)){
  avg_diff_is_if <- round((sum(monthly_is_count$Issued) - sum(monthly_if_count$Transfused))/nrow(monthly_if_count))
} else {
  stop("Check Monthly Counts")
}

```

Column
-------------------------------------------------------

### Transfusions average `r bb_if_monthlyavg` per month

```{r}
monthly_if <- bb_if_data |> 
  group_by(Month, Location = UnitArea) |>  
  count(name = "Count")

g_monthly_if <- ggplot(monthly_if) +
  geom_col(aes(x = Month, y = Count, fill = Location)) +
  labs(title = "Transfused Products by Location") +
  scale_fill_manual(values = mycolors) +
  Graph_Theme

ggplotly(g_monthly_if)

```

### Antibody Screens average `r bb_t_monthlyavg` per month

```{r}
monthly_AS <- bb_t_data |> 
  group_by(Month, Battery) |> 
  count(name = "Count")

g_monthly_AS <- ggplot(monthly_AS) +
  geom_col(aes(x = Month, y = Count, fill = Battery)) +
  labs(title = "Antibody Screen Testing by Battery") +
  scale_fill_manual(values = mycolors) +
  Graph_Theme

ggplotly(g_monthly_AS)

```

Column
---------------------------------------------------------

### Products issued averages `r bb_is_monthlyavg` per month

```{r}
monthly_is <- bb_is_data |> 
  group_by(Month, Components) |> 
  count(name = "Count")

g_monthly_is <- ggplot(monthly_is) +
  geom_col(aes(x = Month, y = Count, fill = Components)) +
  labs(title = "Issued Products by Component Group") +
  scale_fill_manual(values = mycolors) +
  Graph_Theme

ggplotly(g_monthly_is)

```

### Approximately `r avg_diff_is_if` more products are issued per month than is transfused

```{r}
monthly_product_counts <- left_join(monthly_if_count, 
                                    monthly_is_count, 
                                    by = "Month") |> 
  pivot_longer(-c(Month), names_to = "Status", values_to = "Count")

 
g_monthly_product_counts <- ggplot(monthly_product_counts) +
  geom_col(aes(x = Month, y = Count, fill = Status),
           position = "dodge") +
  labs(title = "Blood Product Usage") +
  scale_fill_manual(values = mycolors) +
  Graph_Theme

ggplotly(g_monthly_product_counts)

```

# Turn Around Times

```{r}
xm_ninety <- bb_t_data |> 
  filter(BatTstCode == "XM") |> 
  summarize(test = quantile(TAT, probs = 0.9))

as_ninety <- bb_t_data |> 
  summarize(test = quantile(TAT, probs = 0.9))

stat_ninety <- bb_t_data |> 
  filter(PriorityCodes == "S") |> 
  summarize(test = quantile(TAT, probs = 0.9))

instrument_ninety <- bb_t_data |> 
  filter(ResultMethod != "%BPW") |> ##not manual
  summarize(test = quantile(TAT, probs = 0.9))

##ne1_ninety <- bb_t_data |> 
##  filter(ResultMethod == "NE1") |> ##Immucor Neo Iris
##  summarize(test = quantile(TAT, probs = 0.9))
##
##ec2_ninety <- bb_t_data |> 
##  filter(ResultMethod == "EC2") |> ##Immucor Echo Lumena
##  summarize(test = quantile(TAT, probs = 0.9))

```

Column
-------------------------------------------------------

### Antibody screens are resulted within `r as_ninety` minutes 90% of the time.

```{r}
as_tat_results <- bb_t_data |>  
  group_by(Results, Hour) |> 
  summarize(TAT = quantile(TAT, probs = 0.9)) 

g_as_tat_results <- ggplot(as_tat_results) +
  geom_line(aes(x = Hour, y = TAT, color = Results)) +
  labs(title = "Antibody Screen TAT by Result Category",
       x = "Receipt Hour",
       y = "90th percentile TAT") +
  scale_x_continuous(breaks = seq(1,23,2)) +
  scale_color_manual(values = mycolors) +
  Graph_Theme

ggplotly(g_as_tat_results)

```

### Stat antibody screens are resulted within `r stat_ninety` minutes 90% of the time.

```{r}
as_tat_priority <- bb_t_data |>
  group_by(Priority, Hour) |> 
  summarize(TAT = quantile(TAT, probs = 0.9))

g_as_tat_priority <- ggplot(as_tat_priority) +
  geom_line(aes(x = Hour, y = TAT, color = Priority)) +
  labs(title = "Antibody Screen TAT by Priority",
       x = "Receipt Hour",
       y = "90th percentile TAT") +
  scale_x_continuous(breaks = seq(1,23,2)) +
  scale_color_manual(values = mycolors) +
  Graph_Theme

ggplotly(g_as_tat_priority)

```

Column
-------------------------------------------------------

### Antibody screens from the Instruments are resulted within `r instrument_ninety` minutes 90% of the time.

```{r}
as_tat_method <- bb_t_data |>  
  group_by(Method, Hour) |> 
  summarize(TAT = quantile(TAT, probs = 0.9)) 

g_as_tat_method <- ggplot(as_tat_method) +
  geom_line(aes(x = Hour, y = TAT, color = Method)) +
  labs(title = "Antibody Screen TAT by Result Method",
       x = "Receipt Hour",
       y = "90th percentile TAT") +
  scale_x_continuous(breaks = seq(1,23,2)) +
  scale_color_manual(values = mycolors) +
  Graph_Theme

ggplotly(g_as_tat_method)

```

### Crossmatch antibody screens are resulted within `r xm_ninety` minutes 90% of the time.

```{r}
as_tat_battery <- bb_t_data |>  
  group_by(Battery, Hour) |> 
  summarize(TAT = quantile(TAT, probs = 0.9)) 

g_as_tat_battery <- ggplot(as_tat_battery) +
  geom_line(aes(x = Hour, y = TAT, color = Battery)) +
  labs(title = "Antibody Screen TAT by Battery",
       x = "Receipt Hour",
       y = "90th percentile TAT") +
  scale_x_continuous(breaks = seq(1,23,2)) +
  scale_color_manual(values = mycolors) +
  Graph_Theme

ggplotly(g_as_tat_battery)

```

# Heatmaps

Column
-----------------------------------------------

### Issued Products

```{r}
is_hm <- bb_is_data |> 
  group_by(Weekday, Hour) |> 
  count(name = "Count")

g_is_hm <- ggplot(is_hm, aes(x = Weekday, y = Hour, fill = Count)) +
  geom_tile(lwd = 0.5, linetype = 1, color = "black") +
  scale_fill_distiller(palette = "RdYlGn", direction = -1) + 
  scale_y_continuous(breaks = seq(0,23,1)) +
  geom_text(aes(label = Count), size = 4, color = "black") +
  labs(x = "Day of Week",
       y = "Hour of Day") +
  theme(panel.background = element_blank(),
        axis.ticks = element_blank(),
        legend.position = "none")

ggplotly(g_is_hm)

```


Column
-----------------------------------------------

### Received Antibody Screen Specimens

```{r}
as_hm <- bb_t_data |> 
  group_by(Weekday, Hour) |> 
  count(name = "Count")

g_as_hm <- ggplot(as_hm, aes(x = Weekday, y = Hour, fill = Count)) +
  geom_tile(lwd = 0.5, linetype = 1, color = "black") +
  scale_fill_distiller(palette = "RdYlGn", direction = -1) + 
  scale_y_continuous(breaks = seq(0,23,1)) +
  geom_text(aes(label = Count), size = 4, color = "black") +
  labs(x = "Day of Week",
       y = "Hour of Day") +
  theme(panel.background = element_blank(),
        axis.ticks = element_blank(),
        legend.position = "none")

ggplotly(g_as_hm)

```
