---
title: "Transfusion Service Analysis"
subtitle: "Data Query"
author: "Jessica Jones"
date: "`r format(Sys.Date(),'%B %d, %Y')`"
output: word_document
---

The Laboratory Information System can be queried to pull data for analysis.

# Connection

```{r}
#con <- DBI::dbConnect(odbc::odbc(), dsn = "TEST", timeout = 10)
con <- DBI::dbConnect(odbc::odbc(), dsn = "LAB", timeout = 10)
```

# Transfusions

Each row represents a transfused (%IF) product.
These products are not returned to the Transfusion Service and are billable.

```{sql, connection = con, output.var = "bb_if_data"}

SELECT
    bbi.IssueDate_ODBC
  , mpl_parent.LocCode
  , mpl_parent.LocType
  , bps.UnitArea
  , mct.CompGroup
  , bpsc.CT
FROM 
       SQ.BB_BPSta bps
  JOIN SQ.BB_BPSpcCur bpsc
  ON bps.BB_BPInfo = bpsc.BB_BPInfo
  JOIN SQ.BBMA_CompTy mct
  ON bpsc.CT = mct.CompType
  JOIN SQ.BB_IssuedUnits bbi
  ON bps.BB_BPInfo = bbi.UIDX
  JOIN SQ.MA_PtLoc mpl_specific
  ON bps.ISULoc = mpl_specific.LocIndex
  JOIN SQ.MA_PtLoc mpl_parent
  ON mpl_specific.LocParent = mpl_parent.LocIndex
WHERE
  --between dates
  bbi.IssueDate_ODBC BETWEEN '2021-07-01' AND '2022-06-30'
```

# Issues

Each row represents a product being issued (%IS).
Many locations store products in regulated refrigerators and return the product if not used.  Blood product issue is time consuming and should be considered for staffing even if the product is not ultimately transfused.

```{sql, connection = con, output.var = "bb_is_data"}
SELECT
    bps.UnitArea
  , mct.CompGroup
  , bpsc.CT
  --products only issued once, do not have Activity D/T
  , ISNULL(bpsh.ActivityDate, bpsh.EntryDate) as "IssueDate"
  , ISNULL(bpsh.ActivityTime, bpsh.EntryTime) as "IssueTime"
FROM
       SQ.BB_BPStaHst bpsh
  JOIN SQ.BB_BPSpcCur bpsc
  ON bpsh.BB_BPInfo = bpsc.BB_BPInfo
  JOIN SQ.BBMA_CompTy mct
  ON bpsc.CT = mct.CompType
  LEFT JOIN SQ.BB_BPSta bps
  ON bpsh.BB_BPInfo = bps.BB_BPInfo
WHERE 
  bpsh.Status = '%IS'
  AND (
    bpsh.EntryDate BETWEEN '2021-07-01' AND '2022-06-30'
    OR   
    bpsh.ActivityDate BETWEEN '2021-07-01' AND '2022-06-30'
  )
```

# Testing

Each row represents an antibody screen (%AS) being resulted.
The vast majority of Transfusion Service testing revolved around an antibody screen (%AS).  This test can be utilized for picturing staff and instrument workload.

```{sql, connection = con, output.var = "bb_t_data"}
--%as only
SELECT
    glc.CollectDateODBC
  , glc.CollectTimeODBC
  , glc.ReceiveDateODBC
  , glc.ReceiveTimeODBC
  , glo.BatTstCode
  , glo.PriorityCodes
  , glrh.HistIndex
  , glrh.ResultMethod
  , glrh.ResultTechCodes
  , glrh.ResultLab
  , glrh.CurrentResult
  , glrh.ResultDateODBC
  , glrh.ResultTimeODBC
  , glrh.Result
  , glrh.ResultModifiers
FROM
       SQ.GL_ORDX_MJ base
  JOIN SQ.GL_ORDX glox
  ON base.GL_ORDX_MJ = glox.GL_ORDX_MJ
  JOIN SQ.GL_Specimen gls
  ON glox.LinkToGL_Specimen = gls.RowID
  JOIN SQ.GL_Collect glc
  ON gls.RowID = glc.GL_Specimen
  JOIN SQ.GL_Order glo
  ON glc.RowID = glo.GL_Collect
  JOIN SQ.GL_Test glt
  ON glo.RowID = glt.GL_Order
  JOIN SQ.GL_ResHist glrh
  ON glt.RowID = glrh.GL_Test
  JOIN SQ.MA_OrdCode moc
  ON glo.BatTstCode = moc.OrderCode
WHERE
  --change -# for required look back
  base.ID >= cast(((DateDiff(month,'12/31/1992',GetDate()))-24) as integer)
  --between date range
  AND glc.ReceiveDateODBC BETWEEN '2021-07-01' AND '2022-06-30'
  --and not credited
  AND glo.CreditDate IS NULL
  --and not cancelled
  AND glo.CancelDate IS NULL
  --and not deleted
  AND glo.DeleteDate IS NULL
  --and antibody screen
  AND glt.TestCode = '%AS'
```

# Disconnect

```{r}
DBI::dbDisconnect(con)
```

# Data

The query does not need performed every time so it is stored for use.

```{r}
file_id <- as.numeric(format(Sys.Date(), '%Y'))-1
```

```{r}
saveRDS(bb_if_data, file = paste0("bb_if_data_fy", file_id, ".rds"))
saveRDS(bb_is_data, file = paste0("bb_if_data_fy", file_id, ".rds"))
saveRDS(bb_t_data, file = paste0("bb_if_data_fy", file_id, ".rds"))
```

```{r}
##read data
bb_if_data <- readRDS(file = paste0("../Final/Data/bb_if_data_fy", file_id, ".rds"))
bb_is_data <- readRDS(file = paste0("../Final/Data/bb_is_data_fy", file_id, ".rds"))
bb_t_data <- readRDS(file = paste0("../Final/Data/bb_t_data_fy", file_id, ".rds"))

library(dplyr)

#%IF
str(bb_if_data)
bb_if_data |> group_by(CompGroup) |> count(sort = TRUE)

#%IS
str(bb_is_data)
bb_is_data |> group_by(CompGroup) |> count(sort = TRUE)

#%AS
str(bb_t_data)
bb_t_data |> group_by(ResultMethod) |> count(sort = TRUE)
```

